---
title: "Project 2: STAT302package Tutorial"
author: "Caiwei Tian"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{STAT302package Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Introduction
This package is primarily used for data analysis. It includes four functions to enable computation of t-test, linear model fitting, k nearest neighbor cross-validation and random forest cross-validation.

Use the following command to install the package \texttt{STAT302package}.

```{r,eval=FALSE}
devtools::install_github("caiwet/STAT302package")
```

To begin, load our package first.
```{r setup}
library(STAT302package)
```

# Tutorial for my_t.test
Use \texttt{lifeExp} data from \texttt{my_gapminder}
```{r}
data <- my_gapminder$lifeExp
```

Test 1
Null hypothesis: life expe
```{r}
my_t.test(data, "two.sided", 60)
```

The p-value is greater than 0.05, so the probability of observing a result as or more extreme than what we observed, assuming the null hypothesis is true is greater than the p-value cut-off. Thus, we don't have sufficient evidence to reject the null hypothesis.

Test 2
```{r}
my_t.test(data, "less", 60)
```

This time the p-value is less than 0.05. We can conclude that our results are statistically significant. We are confident to reject null hypothesis and have evidence to support our alternative.

Test 3
```{r}
my_t.test(data, "greater", 60)
```

The p-value is much larger than 0.05. We do not have sufficient evidence to reject null hypothesis.

# Tutorial for my_lm

```{r}
lifeExp <- my_gapminder$lifeExp
gdpPercap <- my_gapminder$gdpPercap
continent <- my_gapminder$continent
input <- data.frame("lifeExp" = lifeExp, "gdpPercap" = gdpPercap, 
                    "continent" = continent)
#my_lm(lifeExp ~ gdpPercap + continent, input)

```





















